---
title: "Week 1 - Homework"
author: "STAT 420, Summer 2021, D. Unger"
date: ''
output:
  html_document: 
    theme: readable
    toc: yes
    highlight: pygments
  pdf_document: default
urlcolor: cyan
---

***

## Exercise 1 (Subsetting and Statistics)

For this exercise, we will use the `msleep` dataset from the `ggplot2` package.

**(a)** Install and load the `ggplot2` package. **Do not** include the installation command in your `.Rmd` file. (If you do it will install the package every time you knit your file.) **Do** include the command to load the package into your environment.

```{r}
library(ggplot2)
msleep <- ggplot2::msleep
```

**(b)** Note that this dataset is technically a `tibble`, not a data frame. How many observations are in this dataset? How many variables? What are the observations in this dataset?

```{r message=FALSE}
num_observations <- nrow(msleep)
num_variables <- ncol(msleep)
description_msleep <- 'The observations show data about sleeping habits of 83 mammals.'
```

- Number of observations in the msleep data set: **`r num_observations`**
- Number of variables: **`r num_variables`**
- What are the observations in this dataset?: **`r description_msleep`**


**(c)** What is the mean hours of REM sleep of individuals in this dataset?

```{r}
(avg_sleep_rem <- mean(msleep$sleep_rem, na.rm = TRUE))
```


**(d)** What is the standard deviation of brain weight of individuals in this dataset?

```{r}
(sd_brainwt <- sd(msleep$brainwt, na.rm = TRUE))
```

**(e)** Which observation (provide the `name`) in this dataset gets the most REM sleep?

```{r}
# Get the row number with the max sleep_rem
max_sleep_rem_row <- which.max(msleep$sleep_rem)
# Get the name column value and output it
max_sleep_rem <- msleep[max_sleep_rem_row, 'name']
max_sleep_rem$name
```

**(f)** What is the average bodyweight of carnivores in this dataset?

```{r}
# Get the bodywt for carnivores
carni_bodywt <- msleep[msleep$vore == 'carni', 'bodywt']
# Calculate the mean
(carni_mean_bodywt <- mean(carni_bodywt$bodywt, na.rm = TRUE))
```


***

## Exercise 2 (Plotting)

For this exercise, we will use the `birthwt` dataset from the `MASS` package.

**(a)** Note that this dataset is a data frame and all of the variables are numeric. How many observations are in this dataset? How many variables? What are the observations in this dataset?

```{r}
library(MASS)
birthwt <- MASS::birthwt
```

```{r message=FALSE}
num_observations <- nrow(birthwt)
num_variables <- ncol(birthwt)
description_birthwt <- 'The observations show risk factors associated with low infant birth weight.'
```

- Number of observations in the birthwt data set: **`r num_observations`**
- Number of variables: **`r num_variables`**
- What are the observations in this dataset?: *`r description_birthwt`*

**(b)** Create a scatter plot of birth weight (y-axis) vs mother's weight before pregnancy (x-axis). Use a non-default color for the points. (Also, be sure to give the plot a title and label the axes appropriately.) Based on the scatter plot, does there seem to be a relationship between the two variables? Briefly explain.

```{r}
plot(bwt ~ lwt, data = birthwt,
        xlab = "Mother's weight before pregnancy (pounds)",
        ylab = "Birth weight (grams)",
        main = "Birth weight vs Mother's weigh",
        pch = 20,
        cex = 2.3,
        col = "lightblue"
)
```

Given how the data points are spread out, does not seem to be a direct relation between these 2 variables. For instance, in the range of 100-150 in the X axis (Mother's weight), the points are spread up and down in the Y axis. Also, It is easy to find cases of very low and very hign Birth Weights given a Mother's weight.

**(c)** Create a scatter plot of birth weight (y-axis) vs mother's age (x-axis). Use a non-default color for the points. (Also, be sure to give the plot a title and label the axes appropriately.) Based on the scatter plot, does there seem to be a relationship between the two variables? Briefly explain.

```{r}
plot(bwt ~ age, data = birthwt,
        xlab = "Mother's age",
        ylab = "Birth weight (grams)",
        main = "Birth weight vs Mother's age",
        pch = 20,
        cex = 2.3,
        col = "lightblue"
)
```

There is no clear tendency of the Birth Weight as the Mother's age increases. 

**(d)** Create side-by-side boxplots for birth weight grouped by smoking status. Use non-default colors for the plot. (Also, be sure to give the plot a title and label the axes appropriately.) Based on the boxplot, does there seem to be a difference in birth weight for mothers who smoked? Briefly explain.

```{r}
bp_conf <- boxplot(bwt ~ smoke, data = birthwt,
        xlab = "Smoking Status (0 = Not smoking, 1 = Smoking)",
        ylab = "Birth weight (grams)",
        main = "Birth weight vs Smoking status",
        pch = 20,
        cex = 2,
        border  = "brown",
        col = "lightblue"
      )
no_smoking_median <- bp_conf$stats[3, 1]
smoking_median <- bp_conf$stats[3, 2]
```

For mothers who smoke, the median (`r smoking_median`) is lower compared to mothers who *do not smoke* (`r no_smoking_median`). Additionally the maximum, the third quartile and the first quartile also drop for mothers who smoke. Given that, there seems to be a relation on birth weight reduction when mothers smoke.

***

## Exercise 3 (Importing Data, More Plotting)

For this exercise we will use the data stored in [`nutrition-2018.csv`](nutrition-2018.csv). It contains the nutritional values per serving size for a large variety of foods as calculated by the USDA in 2018. It is a cleaned version totaling 5956 observations and is current as of April 2018.

The variables in the dataset are:

- `ID` 
- `Desc` - short description of food
- `Water` - in grams
- `Calories` - in kcal
- `Protein` - in grams
- `Fat` - in grams
- `Carbs` - carbohydrates, in grams
- `Fiber` - in grams
- `Sugar` - in grams
- `Calcium` - in milligrams
- `Potassium` - in milligrams
- `Sodium` - in milligrams
- `VitaminC` - vitamin C, in milligrams
- `Chol` - cholesterol, in milligrams
- `Portion` - description of standard serving size used in analysis

```{r message=FALSE}
library(readr)
nutrition <- read_csv("nutrition-2018.csv")
```


**(a)** Create a histogram of `Calories`. Do not modify `R`'s default bin selection. Make the plot presentable. Describe the shape of the histogram. Do you notice anything unusual?

```{r}
hist(nutrition$Calories,
     xlab   = "Calories (kcal)",
     main   = "Histogram of Calories",
     col  = "deepskyblue3",
     border = "gray80"
     )
```

Since the tail of the histogram goes to the right the distribution is skewed right. Something unusual is that the tail of the histogram has a peak for values between 850 and 900.

In other words, the data has a group of considerable size with high Calorie values. I thought it will be interesting to see them:

```{r}
library(knitr)
# Get the row numbers based on the Calories for that peak in the tail
tail_peak_rows <- which(nutrition$Calories >= 850 & nutrition$Calories < 900)
# Get the Description and the Calories columns only
high_calories_obs <- nutrition[tail_peak_rows, c('Desc', 'Calories')]
# Show 10 of them in a table
kable(high_calories_obs[1:10, ], caption = "Food between 850 and 900 Calories")
```


**(b)** Create a scatter plot of calories (y-axis) vs protein (x-axis). Make the plot presentable. Do you notice any trends? Do you think that knowing only the protein content of a food, you could make a good prediction of the calories in the food?

```{r}
plot(Calories ~ Protein, data = nutrition,
        xlab = "Protein (grams)",
        ylab = "Calories (kcal)",
        main = "Calories vs Protein",
        pch = 20,
        cex = 1,
        col = "lightblue"
     )

```

Using only the Calories we won't get a good prediction since matching a line to the data will have a large margin of error. Perhaps getting a higher order feature based on the Calories could give us better predictions.

**(c)** Create a scatter plot of `Calories` (y-axis) vs `4 * Protein + 4 * Carbs + 9 * Fat` (x-axis). Make the plot presentable. You will either need to add a new variable to the data frame, or use the `I()` function in your formula in the call to `plot()`. If you are at all familiar with nutrition, you may realize that this formula calculates the calorie count based on the protein, carbohydrate, and fat values. You'd expect then that the result here is a straight line. Is it? If not, can you think of any reasons why it is not?

```{r}
nutrition$CaloriCount = rep(1, nrow(nutrition))
n <- nutrition

plot(Calories ~ I(4 * Protein + 4 * Carbs + 9 * Fat), 
        data = nutrition,
        xlab = "Calculated Calorie count",
        ylab = "Calories (kcal)",
        main = "Calories vs Calculated Calorie Count",
        pch = 20,
        cex = 1,
        col = "lightblue"
     )
```

The formula used to get the Calorie Count is a good way to predict the Calories. In other words, it will be close to the real Calories observed but in most cases not exactly the same. Different coefficients could also improve the predictions.

***

## Exercise 4 (Writing and Using Functions)

For each of the following parts, use the following vectors:

```{r}
a = 1:10
b = 10:1
c = rep(1, times = 10)
d = 2 ^ (1:10)
```

**(a)** Write a function called `sum_of_squares`.

- Arguments:
    - A vector of numeric data `x`
- Output:
    - The sum of the squares of the elements of the vector $\sum_{i = 1}^n x_i^2$
    
Provide your function, as well as the result of running the following code:

```{r}
sum_of_squares <- function(x) {
  sum(x ^ 2)
}
```

```{r}
sum_of_squares(x = a)
sum_of_squares(x = c(c, d))
```

**(b)** Using only your function `sum_of_squares()`, `mean()`, `sqrt()`, and basic math operations such as `+` and `-`, calculate

\[
\sqrt{\frac{1}{n}\sum_{i = 1}^n (x_i - 0)^{2}}
\]

where the $x$ vector is `d`.

```{r}
# Assigning variables to match the formula (makes it easier to follow)
x <- d
n <- length(x)
result_b <- sqrt(sum_of_squares(x - 0) / n)
```

**Result: `r result_b`**

**(c)** Using only your function `sum_of_squares()`, `mean()`, `sqrt()`, and basic math operations such as `+` and `-`, calculate

\[
\sqrt{\frac{1}{n}\sum_{i = 1}^n (x_i - y_i)^{2}}
\]

where the $x$ vector is `a` and the $y$ vector is `b`.

```{r}
# Assigning variables to match the formula (makes it easier to follow)
x <- a
y <- b
n <- length(x)
result_c <- sqrt(sum_of_squares(x - y) / n)
```

**Result: `r result_c`**

***

## Exercise 5 (More Writing and Using Functions)

For each of the following parts, use the following vectors:

```{r}
set.seed(42)
x = 1:100
y = rnorm(1000)
z = runif(150, min = 0, max = 1)
```

**(a)** Write a function called `list_extreme_values`.

- Arguments:
    - A vector of numeric data `x`
    - A positive constant, `k`, with a default value of `2`
- Output:
    - A list with two elements:
        - `small`, a vector of elements of `x` that are $k$ sample standard deviations less than the sample mean. That is, the observations that are smaller than $\bar{x} - k \cdot s$.
        - `large`, a vector of elements of `x` that are $k$ sample standard deviations greater than the sample mean. That is, the observations that are larger than $\bar{x} + k \cdot s$.

Provide your function, as well as the result of running the following code:

```{r}
list_extreme_values <- function(x, k = 2) {
  x_mean <- mean(x)
  x_sd <- sd(x)
  list(
    small = x[x < x_mean - (k * x_sd)],
    large = x[x > x_mean + (k * x_sd)]
  )
}

```


```{r}
list_extreme_values(x = x, k = 1)
list_extreme_values(x = y, k = 3)
list_extreme_values(x = y, k = 2)
list_extreme_values(x = z, k = 1.5)
```

**(b)** Using only your function `list_extreme_values()`, `mean()`, and basic list operations, calculate the mean of observations that are greater than 1.5 standard deviation above the mean in the vector `y`.

```{r}
mean_large_y = mean(list_extreme_values(x = y, k = 1.5)$large)
```

**Result: `r mean_large_y`**
