---
title: "selection"
author: "AP"
date: "7/14/2021"
output: html_document
---

$$
\text{AIC} = n\log\left(\frac{\text{RSS}}{n}\right) + 2p
$$
$$
\text{BIC} = n\log\left(\frac{\text{RSS}}{n}\right) + \log(n)p
$$



$$
R_a^2 = 1 - \frac{\text{SSE}/(n-p)}{\text{SST}/(n-1)} = 1 - \left(  \frac{n-1}{n-p} \right)(1-R^2)
$$

## Data

```{r, fig.height=8, fig.width=8}
library(faraway)
pairs(seatpos, col = "dodgerblue")
```

```{r}
hipcenter_mod = lm(hipcenter ~ ., data = seatpos)
coef(hipcenter_mod)
```


```{r}
vif(hipcenter_mod)
```

# Exhaustive Search

```{r}
library(leaps)
all_hipcenter_mod = summary(regsubsets(hipcenter ~ ., data = seatpos))
```

```{r}
all_hipcenter_mod
```
Here:

1) is the model with only one predictor with lowest RSS
2) is the model with only two predictors with lowest RSS
etc

```{r}
all_hipcenter_mod$which
```

Among those, the model with the most predictors will have the lowest RSS. So instead of using RSS to decide, we need to penalize the size of the model.

```{r}
all_hipcenter_mod$rss
```

```{r}
all_hipcenter_mod$adjr2
```

When using `Adjusted R^2` we want the max value. 

```{r}
(best_r2_ind <- which.max(all_hipcenter_mod$adjr2))
```

```{r}
all_hipcenter_mod$which[best_r2_ind, ]
```

## AIC

```{r}
p = length(coef(hipcenter_mod))
n = length(resid(hipcenter_mod))
```

```{r}
(hipcenter_mod_aic = n * log(all_hipcenter_mod$rss / n) + 2 * (2:p))
```

With AIC model 3 is also selected.

```{r}
(best_aic_ind = which.min(hipcenter_mod_aic))
all_hipcenter_mod$which[best_aic_ind,]
```

```{r}
hipcenter_mod_best_aic = lm(hipcenter ~ Age + Ht + Leg, data = seatpos)
```

4 here is the number of parameters $p$ or predictors + 1.

```{r}
extractAIC(hipcenter_mod_best_aic)
```


```{r}
plot(hipcenter_mod_aic ~ I(2:p), ylab = "AIC", xlab = "p, number of parameters", 
     pch = 20, col = "dodgerblue", type = "b", cex = 2,
     main = "AIC vs Model Complexity")
```

## BIC

```{r}
(hipcenter_mod_bic = n * log(all_hipcenter_mod$rss / n) + log(n) * (2:p))
```

BIC select the model with one predictor (2 parameters)

```{r}
which.min(hipcenter_mod_bic)
```

```{r}
all_hipcenter_mod$which[1,]
```

```{r}
hipcenter_mod_best_bic = lm(hipcenter ~ Ht, data = seatpos)
```

```{r}
plot(hipcenter_mod_bic ~ I(2:p), ylab = "BIC", xlab = "p, number of parameters", 
     pch = 20, col = "dodgerblue", type = "b", cex = 2,
     main = "BIC vs Model Complexity")
```

# Backward AIC

We provide our full model to the step function. It does the following:

1) Gives RSS and AIC for the full mode <none> entry, and every possible model when removing one of the predictors.

For instance, removing `Ht` gives us the smallest AIC.

2) The next step is the same as above but with `Ht` out of the picture.

3) It stops when removing any of the predictors on a given step, doesn't give a lower AIC.

```{r}
hipcenter_mod_back_aic = step(hipcenter_mod, direction = "backward")
```

The step function returns the model with lowest AIC.

```{r}
coef(hipcenter_mod_back_aic)
```

step uses `extractAIC()` function.

$$
\text{AIC} = n\log\left(\frac{\text{RSS}}{n}\right) + 2p,
$$

```{r}
extractAIC(hipcenter_mod_back_aic)
p <- length(hipcenter_mod_back_aic$coefficients)
n <- nrow(seatpos)

n * log(mean(hipcenter_mod_back_aic$residuals ^ 2)) + 2 * p
```

# Backward BIC

It works the same but now the penalty coefficient that multiplies p is $log(n)$

```{r}
n <- nrow(seatpos)
hipcenter_mod_back_bic = step(hipcenter_mod, direction = "backward", k = log(n))
```
```{r}
coef(hipcenter_mod_back_bic)
```

```{r}
summary(hipcenter_mod)$adj.r.squared
## [1] 0.6000855
summary(hipcenter_mod_back_aic)$adj.r.squared
## [1] 0.6531427
summary(hipcenter_mod_back_bic)$adj.r.squared
## [1] 0.6244149
```

## Forward AIC

When `<none>` reaches the top, adding predictors won't give a smaller AIC.

```{r}

hipcenter_mod_start = lm(hipcenter ~ 1, data = seatpos)
hipcenter_mod_forw_aic = step(
  hipcenter_mod_start, 
  # Max possible model
  scope = hipcenter ~ Age + Weight + HtShoes + Ht + Seated + Arm + Thigh + Leg, 
  direction = "forward")
```

# Forward BIC

```{r}
hipcenter_mod_forw_bic = step(
  hipcenter_mod_start, 
  scope = hipcenter ~ Age + Weight + HtShoes + Ht + Seated + Arm + Thigh + Leg, 
  direction = "forward", k = log(n))
```
## Stepwise AIC

It adds and remove one variable at every step.

Since out initial model only has the Intercept, the first step has nothing to remove.

```{r}
hipcenter_mod_both_aic = step(
  hipcenter_mod_start, 
  scope = hipcenter ~ Age + Weight + HtShoes + Ht + Seated + Arm + Thigh + Leg, 
  direction = "both")
```




