---
title: "logistic"
author: "AP"
date: "7/19/2021"
output: html_document
---

The `noise` follows a binomial distribution.

```{r}
sim_logistic_data = function(sample_size = 25, beta_0 = -2, beta_1 = 3) {
  x = rnorm(n = sample_size)
  eta = beta_0 + beta_1 * x
  p = 1 / (1 + exp(-eta))
  y = rbinom(n = sample_size, size = 1, prob = p)
  data.frame(y, x)
}
```

$$
\log\left(\frac{p({\bf x})}{1 - p({\bf x})}\right) = -2 + 3 x
$$
$$
\begin{aligned}
Y_i \mid {\bf X_i} = {\bf x_i} &\sim \text{Bern}(p_i) \\
p_i &= p({\bf x_i}) = \frac{1}{1 + e^{-\eta({\bf x_i})}} \\
\eta({\bf x_i}) &= -2 + 3 x_i
\end{aligned}
$$
```{r}
set.seed(1)
example_data = sim_logistic_data()
head(example_data)
```
In a lot of ways, lm() is just a more specific version of glm().

By default, glm() uses family = gaussian argument. That is, we’re fitting a GLM with a normally distributed response and the identity function as the link.

```{r}
# ordinary linear regression
fit_lm  = lm(y ~ x, data = example_data)
# logistic regression
fit_glm = glm(y ~ x, data = example_data, family = binomial)
```

```{r}
# more detailed call to glm for logistic regression
fit_glm = glm(y ~ x, data = example_data, family = binomial(link = "logit"))
```

```{r}
fit_glm$coefficients
```

$$
\large
\hat\eta ({\bf x}) = -2.3 + 3.7x
$$
```{r}
predict(fit_glm, newdata = data.frame(x = 1.2), type = "link")
```

$$
\hat{p}({\bf x}) = \frac{e^{\hat{\eta}({\bf x})}}{1 + e^{\hat{\eta}({\bf x})}} = \frac{1}{1 + e^{-\hat{\eta}({\bf x})}}
$$

This is the estimate of the probability that $Y$ is 1 given that $x=1.2$

```{r}
predict(fit_glm, newdata = data.frame(x = 1.2), type = "response")
```

```{r}
plot(y ~ x, data = example_data, 
     pch = 20, ylab = "Estimated Probability", 
     main = "Ordinary vs Logistic Regression")
grid()
abline(fit_lm, col = "darkorange")
curve(predict(fit_glm, data.frame(x), type = "response"), 
      add = TRUE, col = "dodgerblue", lty = 2)
legend("topleft", c("Ordinary", "Logistic", "Data"), lty = c(1, 2, 0), 
       pch = c(NA, NA, 20), lwd = 2, col = c("darkorange", "dodgerblue", "black"))
```

Try different models

```{r}
set.seed(42)
# The intercept seems to have an effect in the proportion of 1s and 0s in the response
# Positive intercept (2) will generate more data points with response 1.
intercept <- 0
# The larger in magnitud the slope is, the sharper the change of probaility from 0 to 1 is.
# Same for negative, in that case it would be decrease of probability is sharper the mode negative the slope is.
slope <- -2

example_data <- sim_logistic_data(sample_size = 50, beta_0 = intercept, beta_1 = slope)

fit_glm <- glm(y ~ x, data = example_data, family = binomial)

fit_glm$coefficients
```

```{r}
plot(y ~ x, data = example_data, 
     pch = 20, ylab = "Probability", 
     main = "True: Orange solid, Estimated: Blue, Dashed")
grid()
curve(predict(fit_glm, data.frame(x), type = "response"), 
      add = TRUE, col = "dodgerblue", lty = 2)

curve(boot::inv.logit(intercept + slope * x), 
      add = TRUE, col = "darkorange", lty = 1)
```

An extreme example.

```{r}
set.seed(42)
# The intercept seems to have an effect in the proportion of 1s and 0s in the response
# Positive intercept (2) will generate more data points with response 1.
intercept <- 10
# The larger in magnitud the slope is, the sharper the change of probaility from 0 to 1 is.
# Same for negative, in that case it would be decrease of probability is sharper the mode negative the slope is.
slope <- -10

example_data <- sim_logistic_data(sample_size = 50, beta_0 = intercept, beta_1 = slope)

fit_glm <- glm(y ~ x, data = example_data, family = binomial)

fit_glm$coefficients
```

```{r}
plot(y ~ x, data = example_data, 
     pch = 20, ylab = "Probability", 
     main = "True: Orange solid, Estimated: Blue, Dashed")
grid()
curve(predict(fit_glm, data.frame(x), type = "response"), 
      add = TRUE, col = "dodgerblue", lty = 2)

curve(boot::inv.logit(intercept + slope * x), 
      add = TRUE, col = "darkorange", lty = 1)
```

```{r}
set.seed(1)
example_data <- sim_logistic_data(sample_size = 50, beta_0 = 10, beta_1 = -10)
fit_glm <- glm(y ~ x, data = example_data, family = binomial)
```

This time estimates look good.

```{r}
fit_glm$coefficients
```

```{r}
set.seed(5)
example_data <- sim_logistic_data(sample_size = 50, beta_0 = 10, beta_1 = -10)
fit_glm <- glm(y ~ x, data = example_data, family = binomial)
```

```{r}
fit_glm$coefficients
```




## SAheart exmaple

```{r}
library(bestglm)
data("SAheart")
```

$$
\log\left(\frac{P[\texttt{chd} = 1]}{1 - P[\texttt{chd} = 1]}\right) = \beta_0 + \beta_{\texttt{ldl}} x_{\texttt{ldl}}
$$

```{r}
chd_mod_ldl = glm(chd ~ ldl, data = SAheart, family = binomial)
plot(jitter(chd, factor = 0.1) ~ ldl, data = SAheart, pch = 20, 
     ylab = "Probability of CHD", xlab = "Low Density Lipoprotein Cholesterol")
grid()
curve(predict(chd_mod_ldl, data.frame(ldl = x), type = "response"), 
      add = TRUE, col = "dodgerblue", lty = 2)
```
In this case we have a z-value instead of a t value.

```{r}
coef(summary(chd_mod_ldl))
```

```{r}
chd_mod_additive = glm(chd ~ ., data = SAheart, family = binomial)
chd_mod_additive$coefficients
```

Our test would be.

$$
H_0: \beta_{\texttt{sbp}} = \beta_{\texttt{tobacco}} = \beta_{\texttt{adiposity}} = \beta_{\texttt{famhist}} = \beta_{\texttt{typea}} = \beta_{\texttt{obesity}} = \beta_{\texttt{alcohol}} = \beta_{\texttt{age}} = 0
$$
This gives the `test statistic`.

```{r}
-2 * as.numeric(logLik(chd_mod_ldl) - logLik(chd_mod_additive))
```

Seems like we prefer the larger model.

```{r}
anova(chd_mod_ldl, chd_mod_additive, test = "LRT")
```

Then we check if all predictors are really needed.

```{r}
chd_mod_selected = step(chd_mod_additive, trace = 0)
coef(chd_mod_selected)
```


No we test for:

$$
H_0: \beta_{\texttt{sbp}} = \beta_{\texttt{adiposity}} = \beta_{\texttt{obesity}} = \beta_{\texttt{alcohol}} = 0
$$
```{r}
anova(chd_mod_selected, chd_mod_additive, test = "LRT")
```

```{r}
summary(chd_mod_selected)
```

### Confident Intervals

`R` uses a method called `profile likelihood`

```{r}
confint(chd_mod_selected, level = 0.99)
```

If we want the Wald Confidence Interval we would do:

$$
\hat{\beta}_j \pm z_{\alpha/2} \cdot \text{SE}[\hat{\beta}_j].
$$

```{r}
confint.default(chd_mod_selected, level = 0.99)
```

### Confidence Intervals for Mean Response

We need to apply the inverse logit function.

$$
\left(\text{logit}^{-1}(\hat{\eta}({\bf x}) - z_{\alpha/2} \cdot \text{SE}[\hat{\eta}({\bf x})] ), \ \text{logit}^{-1}(\hat{\eta}({\bf x}) + z_{\alpha/2} \cdot \text{SE}[\hat{\eta}({\bf x})])\right)
$$

```{r}
new_obs = data.frame(
  sbp = 148.0,
  tobacco = 5,
  ldl = 12,
  adiposity = 31.23,
  famhist = "Present",
  typea = 47,
  obesity = 28.50,
  alcohol = 23.89,
  age = 60
)
```

Tp simply get the prediction value between 0 and 1.

```{r}
predict(chd_mod_selected, new_obs, type = "response")
```


The `predict` function can return the `Standard Error`, $\text{SE}[\hat{\eta}({\bf x})]$

```{r}
eta_hat = predict(chd_mod_selected, new_obs, se.fit = TRUE, type = "link")
eta_hat
```

The fit value is not between 0 and 1, so we need to apply the inverse logit.

```{r}
boot::inv.logit(eta_hat$fit)
```


But we still need an interval.

```{r}
z_crit = round(qnorm(0.975), 2)
eta_hat$fit + c(-1, 1) * z_crit * eta_hat$se.fit
```

But to that we need to apply the logit inverse. Then values are between 0 and 1. This is because we called `predict` with `type = "link"`

```{r}
boot::inv.logit(eta_hat$fit + c(-1, 1) * z_crit * eta_hat$se.fit)
```


### Formula Syntax

All the formula syntax we have been using applies here to. For instance below we add the interaction: `ldl:famhist`

```{r}
chd_mod_interaction = glm(chd ~ alcohol + ldl + famhist + typea + age + ldl:famhist, 
                          data = SAheart, family = binomial)
summary(chd_mod_interaction)
```
Adding a polynomial term.

```{r}
chd_mod_int_quad = glm(chd ~ alcohol + ldl + famhist + typea + age + ldl:famhist + I(ldl^2),
                       data = SAheart, family = binomial)
summary(chd_mod_int_quad)
```

### Deviance

Deviance compares the model to a saturated model. (Without repeated observations, a saturated model is a model that fits perfectly, using a parameter for each observation.) Essentially, deviance is a generalized residual sum of squares for GLMs. Like RSS, deviance decreased as the model complexity increases

```{r}
deviance(chd_mod_ldl)
deviance(chd_mod_selected)
deviance(chd_mod_additive)
```
































