---
title: "NormalDistribution"
author: "AP"
date: "5/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

IQ scores follow a Normal Distribution with mean 100 and standard deviation 15

$$
X \sim N(\mu = 100, \sigma^2 = 15^2)
$$

What is the probability of person having IQ below 115
We get this with CDF, Cumulative Dist Function.
Basically is the cummulative probability from -infinity to 115

p in pnorm stands for Probability

```{r}
(pnorm(115, mean = 100, sd = 15))
```

Standarizing a variable
If we do this to variable X we get the Standard Norm

$$
Z = \frac{X - \mu}{\sigma} \sim N(0, 1)
$$

Where is 115 located with respect to the mean?
Answer, 115 if 1 standard deviation from the mean


```{r}
(115 - 100) / 15
?pnorm
```


By default pnorm uses mean 0 and sd = 1, so pnorm(1) gives the same value

```{r}
pnorm(1)
```

What is the height of the density curve at an IQ of 115?
This is not cumulative, just the density (% of samples) at exactly 115

```{r}
dnorm(115, mean = 100, sd = 15)
# At 85 is the same because 85 and 115 are 1 sd from the mean
dnorm(85, mean = 100, sd = 15)

```
Probability of a person to have IQ of 115 or less
Again, note how this is cumulative probability from -inf to 115

```{r}
pnorm(115, mean = 100, sd = 15)
```



Probability of a person to have IQ between 100 and 115
Note that 0.3413447 * 2 is density within 1 standard deviation which matches with
the rule of 68% of density being 1 sd from the mean in a norm dist

```{r}
pnorm(115, mean = 100, sd = 15) - pnorm(100, mean = 100, sd = 15)
```

R norm functions are vectorized

```{r}
pnorm(c(100, 115), mean = 100, sd = 15)
diff(pnorm(c(100, 115), mean = 100, sd = 15))
```

Probability of a person to have IQ > 130

```{r}
1 - pnorm(130, mean = 100, sd = 15)
pnorm(130, mean = 100, sd = 15, lower.tail = FALSE)
```

What IQ is needed to be in the top 5%?

$$
P[X > c] = 0.05
$$

qnorm is the opposite of pnorm in the sense that we get X given a probability p

```{r}
# The IQ needed to be in the lower 5% (not very smart people :P)
qnorm(0.05, mean = 100, sd = 15)
?qnorm
# If get the tail on the right we have the IQ needed to be in the top 5% (smarter people... so to speak) 
qnorm(0.05, mean = 100, sd = 15, lower.tail = FALSE)
# Another way to do it is to get the IQ where the Probability density reaches 95%
qnorm(0.95, mean = 100, sd = 15)
```

What is the probability that someone has an IQ more than 2 sd from the mean?

```{r}
left_side <- pnorm((100 - (2 * 15)), mean = 100, sd = 15)
right_side <- pnorm((100 + (2 * 15)), mean = 100, sd = 15, lower.tail = FALSE)
left_side + right_side

# A smarter way to do it
1 - diff(pnorm(c(70, 130), mean = 100, sd = 15))
# Since the distribution is symmetrical, this also works 
pnorm(70, mean = 100, sd = 15) * 2
```

Note the we get the same value from above with defaults for sd and mean (1, 0),
which is the standard normal distribution

```{r}
# Values located more than -2 * sd from the mean
pnorm(-2) * 2
```

Get 20 random samples from the distribution

```{r}
rnorm(20, mean = 100, sd = 15)
```

# Vectorization

We can get many values from different distributions at the same time

```{r}
pnorm(c(0.5, 1, 0), mean = c(-1, 0, 1), sd = c(2, 1, 0.5))

pnorm(0, mean = c(-1, 0, 1), sd = c(2, 1, 0.5))
```

## Other Continuous distributions

```{r}
dexp(2, rate=0.5)

```









