---
title: "quiz-week2"
author: "AP"
date: "5/27/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Week 2 Quiz

# Question 1

Consider the simple linear regression model 

$$Y = 10 + 5x + \epsilon$$

where 

$$\epsilon \sim N(0, \sigma^2 = 16)$$

Calculate the probability that Y is less than 6 given that x = 0.
Again, 68% of the prob density is in the range of 1 sd from the mean, here sd is 4

```{r}
pnorm(4, mean = 0, sd = 4) - pnorm(-4, mean = 0, sd = 4)
```
```{r}
pnorm(6, mean = 10, sd = 4)
```

# Question 2

Using the SLR model in Question 1, what is the probability that Y is greater than 3 given that x = -1?

```{r}
(expected_y <- 10 + 5*(-1))
pnorm(3, mean = expected_y, sd = 4, lower.tail = FALSE)
```

# Question 3

Using the SLR model in Question 1, what is the probability that YY is greater than 3 given that x = -2?

```{r}
(expected_y <- 10 + 5*(-2))
pnorm(3, mean = expected_y, sd = 4, lower.tail = FALSE)
```
# Question 4

Suppose we would like to predict the duration of an eruption of the Old Faithful geyser in Yellowstone National Park based on the waiting time before an eruption. Fit a simple linear model in R that accomplishes this task

What is the estimate of the intercept parameter?

```{r}
?faithful
f_fit <- lm(eruptions ~ waiting, data = faithful)
f_fit$coefficients[1]
```

# Question 5

What is the estimate of the slope parameter?  

```{r}
f_fit$coefficients[2]
plot(eruptions ~ waiting, data = faithful)
abline(f_fit)
```

# Question 6

Use the fitted model to predict the duration of an eruption based on a waiting time of 80 minutes.

```{r}
predict(f_fit, newdata = data.frame(waiting = 80))
```

# Question 7

Use the fitted model to predict the duration of an eruption based on a waiting time of 120 minutes.

```{r}
predict(f_fit, newdata = data.frame(waiting = 120))
```

# Question 8

Of the predictions that you made for 80 and 120 minutes, which is more reliable?

```{r}
# 80 is more reliable because 120 is out of range of the observed data
max(faithful$waiting)
min(faithful$waiting)
faithful[faithful$waiting == 80, ]
```

# Question 9

Calculate the RSS for the fitted model. 
RSS means Residual Sum of Squares

```{r}
y <- faithful$eruptions
y_hat <- predict(f_fit, newdata = faithful)
RSS <- sum((y - y_hat) ^ 2)
```

# Question 10

What proportion of the variation in eruption duration is explained by the linear relationship with waiting time?

```{r}
TSS <- sum((y - mean(y)) ^ 2)
RegSS <- sum((y_hat - mean(y)) ^ 2)
RegSS / TSS
```

# Question 11

Calculate the standard deviation of the residuals of the fitted model.

```{r}
summary(f_fit)
names(f_fit)
sd(f_fit$residuals)

sd(y - y_hat)
```

