---
title: "mlr_intervals"
author: "AP"
date: "6/9/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Intervals and Test for MLR

## Importing Data

```{r}
autompg <- read.table(
  "http://archive.ics.uci.edu/ml/machine-learning-databases/auto-mpg/auto-mpg.data",
  stringsAsFactors = FALSE
)
#give headers
colnames(autompg) <- c("mpg","cyl","disp","hp","wt","acc","year","origin","name")
# remove missing data marked as ?
autompg <- subset(autompg, autompg$hp != "?")
# remove a record that has issues
autompg <- subset(autompg, autompg$name != "plymouth reliant")
# give the dataset row names, based on the engine, year and name
rownames(autompg) <- paste(autompg$cyl, "cylinder", autompg$year, autompg$name)
# remove the name and origin columns
autompg <- subset(autompg, select = c("mpg","cyl","disp","hp","wt","acc","year"))
# change horsepower to numeric
autompg$hp <- as.numeric(autompg$hp)

str(autompg)
```

## Fitting MLR

$$
Y_i = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} + \epsilon_i
$$

```{r}
mpg_model <- lm(mpg ~ wt + year, data = autompg)
mpg_model$coefficients
summary(mpg_model)
```

# Coefficients

Now the mode contains the same coef as in did for SLR but know
with one more row since we fit for wt and yeat.

The coefficients for wt represent values when **the other predictors are also
in the model**

```{r}
summary(mpg_model)$coef
```

Here for instance is the p value when we are testing for year

```{r}
summary(mpg_model)$coef["year", "Pr(>|t|)"]
```

# Confidence Intervals

We can get a Confidence Interval in the same way since functions are `vectorized`

```{r}
confint(mpg_model, level = 0.99)
```

From the CI above and the p value for `year`, we would reject $H_0: \beta_2 = 0$

# CI for new observations

```{r}
(new_cars <- data.frame(wt = c(3500, 5000), year = c(76, 81)))
```

```{r}
predict(mpg_model, newdata = new_cars,
        interval = "confidence", level = 0.99)
```

# PI for new observations

```{r}
predict(mpg_model, newdata = new_cars,
        interval = "prediction", level = 0.99)
```


# Hidden extrapolations

Individually, seems like the new values are in the rages of their respective predictors.

```{r}
library("dplyr")
between(c(3500, 5000), min(autompg$wt), max(autompg$wt))
between(c(76, 81), min(autompg$year), max(autompg$year))
```

But combined, one of them is an outlier:

The data used to fit the model does not have cases similar to ~5000 lbs and 1980s

```{r}
plot(year ~ wt, data = autompg, pch = 20, col = "dodgerblue", cex = 1.5)
points(new_cars, col = "darkorange", cex = 2.5, pch = "X")
```


# Critical values

```{r}
confint(mpg_model, level = 0.99, param = "wt")
summary(mpg_model)$coef
```

```{r}
beta_1_hat <- summary(mpg_model)$coef["wt", "Estimate"]
es_err_wt <- summary(mpg_model)$coef["wt", "Std. Error"]

# Get the degrees of freedom in 2 ways
nrow(autompg) - length(mpg_model$coef)
df <- summary(mpg_model)$df[2]

abs(qt((1 - 0.99) / 2, df = df))
(crit <- qt((1 - 0.99) / 2, df = df, lower.tail = FALSE))
```

Now we can calculate margins for `wt`

```{r}
c(beta_1_hat - crit * es_err_wt, beta_1_hat + crit * es_err_wt)
```

