---
title: "nested_models"
author: "AP"
date: "6/9/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Nested Models

## Importing Data

```{r}
autompg <- read.table(
  "http://archive.ics.uci.edu/ml/machine-learning-databases/auto-mpg/auto-mpg.data",
  stringsAsFactors = FALSE
)
#give headers
colnames(autompg) <- c("mpg","cyl","disp","hp","wt","acc","year","origin","name")
# remove missing data marked as ?
autompg <- subset(autompg, autompg$hp != "?")
# remove a record that has issues
autompg <- subset(autompg, autompg$name != "plymouth reliant")
# give the dataset row names, based on the engine, year and name
rownames(autompg) <- paste(autompg$cyl, "cylinder", autompg$year, autompg$name)
# remove the name and origin columns
autompg <- subset(autompg, select = c("mpg","cyl","disp","hp","wt","acc","year"))
# change horsepower to numeric
autompg$hp <- as.numeric(autompg$hp)

str(autompg)
```

# Significance of Regression

The null model has both $\hat\beta_1$ and $\hat\beta_2$ as `0`

```{r}
null_mpg_model <- lm(mpg ~ 1, data = autompg)
full_mpg_model <- lm(mpg ~ wt + year, data = autompg)
(nf_anova <- anova(null_mpg_model, full_mpg_model))
```

RSS for the null model is just the Sum Of Squared Residuals of the response.

This is the SST because uses the residual between y and its mean.

```{r}
nf_anova[1, "RSS"]
(sst <- sum((autompg$mpg - mean(autompg$mpg)) ^ 2))

```

Then SSReg is a portion of SST

```{r}
nf_anova[2, "Sum of Sq"]
(ssreg <- sum((full_mpg_model$fitted.values - mean(autompg$mpg)) ^ 2))
```

The difference between those 2 is the SSE

```{r}
nf_anova[1, "RSS"] - nf_anova[2, "Sum of Sq"]
sum((autompg$mpg - full_mpg_model$fitted.values) ^ 2)
```

Note that the `ful_model` and `anova` have the same, p value, F statistic and degrees of freedom.

Also from the model we can see small p values for both `wt` and `year`, so it seems together they turned out to be very significant. **But that doesn't mean they are individually significant.**

```{r}
summary(full_mpg_model)
nf_anova
```


If we check R^2 we can tell how well the model is explaining the data.

$R^2$

```{r}
summary(full_mpg_model)$r.squared
ssreg / sst
```

# Nested Models

```{r}
null_mpg_model <- lm(mpg ~ wt + year, data = autompg)
full_mpg_model <- lm(mpg ~ ., data = autompg)
(nf_anova <- anova(null_mpg_model, full_mpg_model))
```

In this case the full model has a large `p value` due to a small `F statistic` value. The $H_0$ is defined as:

$$
H_0: \beta_{cyl} = \beta_{disp} = \beta_{hp} = \beta_{acc} = 0
$$

With a reasonable $\alpha$ we will `FTR the null Hypothesis` meaning that **those predictors are not significant when `wt` and `year` are in the model.**


# Calculating p values manually

We use the pf function and pass the F statistic

```{r}
nf_anova[2, "F"]
# p - q degrees of freedom From the diff of sum of squares 
(p_q <- nf_anova[2, "Df"])
(n_p <- summary(full_mpg_model)$df[2])

pf(nf_anova[2, "F"], df1 = p_q, df2 = n_p, lower.tail = FALSE)
nf_anova[2, "Pr(>F)"]
```


# Calculating F statistic manually

Here we decompose the error made by the `null model`, not the variance of `y`

```{r}
(diff_sum_sq <- sum((full_mpg_model$fitted.values - null_mpg_model$fitted.values) ^ 2))
nf_anova[2, "Sum of Sq"]
nf_anova[1, "RSS"] - nf_anova[2, "RSS"] 


(full_sum_sq <- sum((autompg$mpg - full_mpg_model$fitted.values) ^ 2))
nf_anova[2, "RSS"]
nf_anova[1, "RSS"] - nf_anova[2, "Sum of Sq"]

(F_sta <- (diff_sum_sq / p_q) / (full_sum_sq / n_p))
nf_anova[2, "F"]
```


# t vs F

```{r}
null_mpg_model <- lm(mpg ~ wt + year, data = autompg)
full_mpg_model <- lm(mpg ~ wt + year + acc, data = autompg)
(nf_anova <- anova(null_mpg_model, full_mpg_model))
```

Note that the `t value` ^ 2 for `acc` in the full model is the `F statistic` in the anova table

```{r}
summary(full_mpg_model)$coef
summary(full_mpg_model)$coef["acc", 3] ^ 2
nf_anova[2, "F"]

```

If we do this manually note that p - q becomes 1. The null model has only one less predictor than the full model has.

```{r}
(p_q <- nf_anova[2, "Df"])
(n_p <- summary(full_mpg_model)$df[2])
(diff_sum_sq <- sum((full_mpg_model$fitted.values - null_mpg_model$fitted.values) ^ 2))
(full_sum_sq <- sum((autompg$mpg - full_mpg_model$fitted.values) ^ 2))
(F_sta <- (diff_sum_sq / p_q) / (full_sum_sq / n_p))
```

```{r}
(beta_3_hat <- full_mpg_model$coefficients[4])
Sxx <- sum((autompg$acc - mean(autompg$acc)) ^ 2)

# Residual stantard error
(s_e <- summary(full_mpg_model)$sigma)
sqrt(sum((autompg$mpg - full_mpg_model$fitted.values)^2) / n_p)

# How do I calculate the Std. Error correctly? are Sxx and s_e correct?
(s_err <- summary(full_mpg_model)$coef["acc", "Std. Error"])
(manual_beta_3_hat_t <- (beta_3_hat - 0) / (s_e / sqrt(Sxx)))

(manual_beta_3_hat_t <- (beta_3_hat - 0) / s_err)
summary(full_mpg_model)$coef["acc", "t value"]
```




































