---
title: "quiz-week4"
author: "AP"
date: "6/10/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Question 1

Consider testing for significance of regression in a multiple linear regression model with 9 predictors and 30 observations. If the value of the $F$ test statistic is 2.4, what is the p-value of this test?

```{r}
1 - pf(2.4, df1 = 10 - 1, df2 = 30 - 10)
```

# Question 2

What is the p-value for testing $H_0: \beta_1 = 0 \ \text{vs} \ H_1: \beta_1 \neq 0$
in a multiple linear regression model with 5 predictors and 20 observations if the value of the t test statistic is -1.3? 

```{r}
2 * pt(-abs(-1.3), df = 20 - 6)

1 - pf((-1.3)^2, df1 = 6 - 5, df2 = 20 - 6)
```

# Question 3

Consider the true model

$$Y = 3 + 2 x_1  + 0.5 x_2  + 5 x_3 + \epsilon$$

where 

$$ϵ∼N(0,σ^2 = 9) $$

What is $\text{SD}[\hat{\beta}_2]$ given the values of predictors above?


$$\text{Var}[\hat{\beta}] = \sigma^2 \left(  X^\top X  \right)^{-1}$$

$$\text{Var}[\hat{\beta}_j] = \sigma^2 C_{jj}$$

```{r}
set.seed(42)
x_values = data.frame(
  x1 = runif(15),
  x2 = runif(15),
  x3 = runif(15)
)

X <- cbind(1, x_values$x1, x_values$x2, x_values$x3)
C <- solve(t(X) %*% X)

sqrt(9 * C[3, 3])

```

# Question 4

For Questions 4-11, use the swiss dataset, which is built into R.

Fit a multiple linear regression model with Fertility as the response and the remaining variables as predictors. You should use `?swiss` to learn about the background of this dataset.

Use your fitted model to make a prediction for a Swiss province in 1888 with:

- 54% of males involved in agriculture as occupation
- 23% of draftees receiving highest mark on army examination
- 13% of draftees obtaining education beyond primary school
- 60% of the population identifying as Catholic
- 24% of live births that live less than a year

```{r}
#?swiss
full_model <- lm(Fertility ~ . , data = swiss)

new_fert <- data.frame(
  Agriculture = 54,
  Examination = 23,
  Education = 13,
  Catholic = 60,
  Infant.Mortality = 24
)
predict(full_model, newdata = new_fert)

bh <- full_model$coef
bh[1] + bh[2]*54 + bh[3]*23 + bh[4]*13 + bh[5]*60 + bh[6]*24
```


# Question 5

Create a 99% confidence interval for the coefficient for Catholic. Report the upper bound of this interval.

```{r}
confint(full_model, level = 0.99, parm = "Catholic")[2]
```

# Question 6

Calculate the p-value of the test 

$$H_0: \beta_{\text{Examination}} = 0 \ \text{vs} \ H_1: \beta_{\text{Examination}} \neq 0$$
 
```{r}
summary(full_model)$coef["Examination", "Pr(>|t|)"]
```

# Question 7

Create a 95% confidence interval for the average Fertility for a Swiss province in 1888 with:

- 40% of males involved in agriculture as occupation
- 28% of draftees receiving highest mark on army examination
- 10% of draftees obtaining education beyond primary school
- 42% of the population identifying as Catholic
- 27% of live births that live less than a year

Report the lower bound of this interval.

```{r}
new_fert <- data.frame(
  Agriculture = 40,
  Examination = 28,
  Education = 10,
  Catholic = 42,
  Infant.Mortality = 27
)
predict(full_model, newdata = new_fert,
        interval = "confidence", level = 0.95)[1, "lwr"]
```


# Question 8

Create a 95% prediction interval for the average Fertility for a Swiss province in 1888 with:

- 40% of males involved in agriculture as occupation
- 28% of draftees receiving highest mark on army examination
- 10% of draftees obtaining education beyond primary school
- 42% of the population identifying as Catholic
- 27% of live births that live less than a year

Report the lower bound of this interval.

```{r}
new_fert <- data.frame(
  Agriculture = 40,
  Examination = 28,
  Education = 10,
  Catholic = 42,
  Infant.Mortality = 27
)
predict(full_model, newdata = new_fert,
        interval = "prediction", level = 0.95)[1, "lwr"]
```

# Question 9

Report the value of the $F$ statistic for the significance of regression test.

```{r}
f_val <- summary(full_model)$fstatistic["value"]
```

# Question 10

Carry out the significance of regression test using $\alpha = 0.01$. What decision do you make?

```{r}
1 - pf(f_val, df1 = 6 - 1, df2 = nrow(swiss) - 6)
p_val <- pf(f_val, df1 = 6 - 1, df2 = nrow(swiss) - 6, lower.tail = FALSE)

# At least one predictor is explaining 70% of the variance
summary(full_model)$r.squared

stat_decide <- function(p_value, alpha = 0.01) {
  ifelse(p_value < alpha, "Reject the Null Hypothesis", "FTR the Null Hypothesis")
}

stat_decide(p_val, alpha = 0.01)
```


# Question 11

Consider a model that only uses the predictors Education, Catholic, and Infant.Mortality. Use an FF test to compare this with the model that uses all predictors. Report the p-value of this test.

```{r}
full_model <- lm(Fertility ~ . , data = swiss)
null_model <- lm(Fertility ~ Education + Catholic + Infant.Mortality, data = swiss)

an <- anova(null_model, full_model)

an[2, "Pr(>F)"]
```


# Question 13

```{r}
full_model <- lm(Employed ~ GNP + Population, data = longley)
null_model1 <- lm(Employed ~ GNP, data = longley)
null_model2 <- lm(Employed ~ Population, data = longley)

anova(null_model1, full_model)
anova(null_model2, full_model)
full_model$coef
```


# Question 14

Suppose you have an SLR model for predicting IQ from height. The estimated coefficient for height is positive. Now, we add a predictor for age to create an MLR model. After fitting this new model, the estimated coefficient for height must be:

```{r}
model1 <- lm(Employed ~ GNP, data = longley)
model2 <- lm(Employed ~ GNP + Population, data = longley)

model1$coef
model2$coef
```























