---
title: "quiz-week3"
author: "AP"
date: "6/3/2021"
output: html_document
---

# Question 1

Consider a random variable XX that has a tt distribution with 55 degrees of freedom. Calculate $P[|X| > 2.1]$ .

```{r}
pt(2.1, df = 5, lower.tail = FALSE) # Wrong
pt(2.1, df = 5, lower.tail = FALSE) * 2
```

# Question 2

Calculate the critical value used for a 90% confidence interval about the slope parameter of a simple linear regression model that is fit to 10 observations. (Your answer should be a positive value.)  

```{r}
(1 - 0.9) / 2
qt(0.05, df = 8, lower.tail = FALSE)
```

# Question 3

Consider the true simple linear regression model  

$Y_i = 5 + 4 x_i + \epsilon_i \qquad \epsilon_i \sim N(0, \sigma^2 = 4) \qquad i = 1, 2, \ldots 20$

  Given $S_{xx} = 1.5$, calculate the probability of observing data according to this model, fitting the SLR model, and obtaining an estimate of the slope parameter greater than 4.2. 

In other words, calculate $P[\hat{\beta}_1 > 4.2]$
 
```{r}
Sxx <- 1.5
sigma_2 <- 4
beta_1 <- 4
var_beta_1_hat <- sigma_2 / Sxx
1 - pnorm(4.2, mean = beta_1, sd = sqrt(var_beta_1_hat))
```

# Question 4

Suppose we would like to predict the duration of an eruption of the Old Faithful geyser in Yellowstone National Park based on the waiting time before an eruption. Fit a simple linear model in R that accomplishes this task.

What is the value of $\text{SE}[\hat{\beta}_1]$

```{r}
?faithful
f_model <- lm(eruptions ~ waiting, data = faithful)
coef(summary(f_model))
coef(summary(f_model))["waiting", "Std. Error"]
```

# Question 5

What is the value of the test statistic for testing $H_0: \beta_0 = 0$ vs $H_1: \beta_0 \neq 0$ 


```{r}
coef(summary(f_model))
coef(summary(f_model))[1, 3]
```

# Question 6

What is the value of the test statistic for testing $H_0: \beta_1 = 0$ vs $H_1: \beta_1 \neq 0$  ?

```{r}
coef(summary(f_model))[2, 3]
```

# Question 7

```{r}
alpha <- 0.01
f_model$coefficients
ifelse(coef(summary(f_model))[2, 4] < alpha, "reject", "FTR")
```

# Question 8

Calculate a 90% confidence interval for $\beta_0$. Report the upper bound of this interval.

```{r}
confint(f_model, level = 0.9)[1, 2]
```

# Question 9

Calculate a 95% confidence interval for $\beta_1$. Report the length of the margin of this interval.

```{r}
conf_inter <- confint(f_model, level = 0.95)
conf_inter[2, 2] - conf_inter[2, 1] # wrong
```

```{r}
conf_inter <- confint(f_model, level = 0.95)

beta_1_hat_se <- summary(f_model)$coefficients[2, 2]
(1 - 0.95) / 2
crit <- qt(0.025, df = nrow(faithful) - 2, lower.tail = FALSE)
(crit * beta_1_hat_se)

beta_1_hat <- f_model$coefficients[2]
beta_1_hat - (crit * beta_1_hat_se)
beta_1_hat + (crit * beta_1_hat_se)

```

# Question 10

Create a 90% confidence interval for the mean eruption duration for a waiting time of 81 minutes. Report the lower bound of this interval.  

```{r}

conf_interv <- predict(f_model, newdata = data.frame(waiting = 81),
        interval = c("confidence"), level = 0.9)
conf_interv[1, "lwr"]
```

# Question 11

Create a 99% prediction interval for a new observation's eruption duration for a waiting time of 72 minutes. Report the upper bound of this interval.  

```{r}
pred_interv <- predict(f_model, newdata = data.frame(waiting = 72),
        interval = c("prediction"), level = 0.99)
pred_interv[1, "upr"]
```

# Question 12

Consider a 90% confidence interval for the mean response and a 90% prediction interval, both at the same xx value. Which interval is narrower?  

```{r}
predict(f_model, newdata = data.frame(waiting = 81),
        interval = c("confidence"), level = 0.9)
predict(f_model, newdata = data.frame(waiting = 83),
        interval = c("prediction"), level = 0.9)
```

























